\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-10-18 reviews/response.Rmd

> Par 2. rephrase "more easily pop-up to viewers".

We've rephrased to "to accentuate unusual patterns, such as those related to special events, for viewers".

> Page 7. Second line. You might want to add that fifth row could be blank. Wrapping around the last few days could be confusing. Could this be made optional?

Thanks, we added the note "The fifth row could be blank for February if the month starts on Monday." in the paragraph. We have mentioned the option of 6-week format as part of the future work.

@@

We've reduced the size of panel labels to give more room to the data display. Time labels are now made to show 6 and 18 to avoid confusion.

The colour scheme is suggested by reviewer 3 before, which is colour-blind and grey-scale print friendly.

> Page 23. Suggest dropping figure 13. Or move to supplementary material.

We dropped the boxplots as in figure 13, and the corresponding paragraph.

@@

TODO

-->

> p17 Figure 10  Why not align by days of the week?

We have changed figure 10 (now figure 8) to align by days of the week instead of by days of the month.

@@

> p22 "with noticeable variations on Thursdays" based on only 26 data points is over-interpretation.

It's half-hourly data, meaning 26 Thursdays * 48 per day = 1248 data points.

> p22 "The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns."  The vertical scaling makes it very difficult to see this.  Inspecting the data with individual scales offers some support but not a lot.

We dropped the boxplot as in figure 13, and the corresponding paragraph including this sentence.

### 2019-10-18 calendar-vis.Rmd
change lag scatterplots to using weekly display

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
```

@@

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)). Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes. An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses modular arithmetic, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to weeks of the month and days of the week respectively. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The notation for creating these cells is as follows:

@@

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ is the row and $1 \le n \le N$ is the column; $b$ denotes the small amount of white space between each month for visual separation. Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.

@@

Note that for the vertical direction, the top left is the starting point of the grid (in Figure \@ref(fig:month-diagram)), which is easier to lay out and why the subtraction is performed. Within each cell, the starting position is the bottom left.

@@

```{r scatterplot, fig.height = 11, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
```

(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the daily calendar layout. Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days (blue) than work days (red).

@@

@@

```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households. Suggested by the medians, household 3 uses more energy than the others on the weekdays. By contrast, household 2 sees considerably larger variability."}
```

Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend. Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between weekdays and weekends. As for household 4, their home routine starts after 18 on weekdays. Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.

>>>>

```{r ped-map, fig.width = 7, out.width = "70%", fig.cap = "(ref:ped-map-cap)"}
```

@@

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), variations of its usage (Section \@ref(sec:variations)), and the full faceting extension equipped with formal labels and axes (Section \@ref(sec:facet-calendar)). An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses modular arithmetic, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to weeks of the month and days of the week respectively. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The fifth row could be blank for February if the month starts on Monday. The notation for creating these cells is as follows:

@@

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ is the row and $1 \le n \le N$ is the column; $b$ denotes the small amount of white space between each month for visual separation.

@@

Note that for the vertical direction, the top left is the starting point of the grid, which is easier to lay out and why the subtraction is performed. Within each cell, the starting position is the bottom left.

@@

```{r scatterplot, fig.height = 12, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
```

(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the weekly calendar layout. Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days (blue) than work days (red).

@@

<!--

@@

```{r dow, fig.height = 7, fig.cap = "(ref:dow-cap)"}
```

(ref:dow-cap) Examining daily variability of household's energy usage by day of week, using boxplots. The medians suggests that household 3 uses more energy than the others on the weekdays. Household 2 sees considerably larger variability given the scales. However, the sub-daily variations cannot been seen from the plot.

-->

Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend. Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between weekdays and weekends. As for household 4, their home routine starts after 18 on weekdays. Figures \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

(ref:hod-cap) Examining sub-daily variability of energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.

### 2019-10-18 calendar-vis.tex
change lag scatterplots to using weekly display

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), and variations of its usage (Section \ref{sec:variations}). Section \ref{sec:facet-calendar} explains the full faceting extension, that is equipped with formal labels and axes. An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

@@

To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) is the row and \(1 \le n \le N\) is the column; \(b\) denotes the small amount of white space between each month for visual separation. Figure \ref{fig:year-diagram} illustrates this layout where \(M = 3\) and \(N = 4\).

@@

Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}), which is easier to lay out and why the subtraction is performed. Within each cell, the starting position is the bottom left.

@@

\caption{Lag scatterplots, with local scaling, in the daily calendar layout. Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days (blue) than work days (red).}\label{fig:scatterplot}
\end{figure}

@@

Figure \ref{fig:dow} shows the energy use across days of week using boxplots. Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the weekdays than other households. The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Fridays, and much higher usage over the weekends. The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the weekdays, reflecting work and leisure patterns.

{\centering \includegraphics[width=\textwidth]{figure/dow-1}

\caption{Boxplots of daily energy usage against day of week for four households. Suggested by the medians, household 3 uses more energy than the others on the weekdays. By contrast, household 2 sees considerably larger variability.}\label{fig:dow}
\end{figure}

Figure \ref{fig:hod} shows energy consumption against time of day, separately by weekday and weekend. Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between weekdays and weekends. As for household 4, their home routine starts after 18 on weekdays. Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

\caption{Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.}\label{fig:hod}
\end{figure}

>>>>

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), variations of its usage (Section \ref{sec:variations}), and the full faceting extension equipped with formal labels and axes (Section \ref{sec:facet-calendar}). An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

@@

To create the layout for a full year, \((m, n)\) denotes the position of the month arranged in the plot, where \(1 \le m \le M\) is the row and \(1 \le n \le N\) is the column; \(b\) denotes the small amount of white space between each month for visual separation.

@@

Note that for the vertical direction, the top left is the starting point of the grid, which is easier to lay out and why the subtraction is performed. Within each cell, the starting position is the bottom left.

@@

\caption{Lag scatterplots, with local scaling, in the weekly calendar layout. Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days (blue) than work days (red).}\label{fig:scatterplot}
\end{figure}

@@

Figure \ref{fig:hod} shows energy consumption against time of day, separately by weekday and weekend. Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between weekdays and weekends. As for household 4, their home routine starts after 18 on weekdays. Figures \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

\caption{Examining sub-daily variability of energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.}\label{fig:hod}
\end{figure}
